Predictions

### 2021-04-10 analysis/04_predictions.Rmd
Do prediction stuff with latent human rights

  # Load actual model objects
  tar_load(c(m_pts_baseline_train, m_pts_v2csreprss_train,
             m_pts_total_train, m_pts_advocacy_train, 
             m_pts_entry_train, m_pts_funding_train,
             m_pts_baseline_rewb_train, m_pts_v2csreprss_rewb_train, 
             m_clphy_baseline_train, m_clphy_v2csreprss_train, 
             m_clphy_total_train, m_clphy_advocacy_train, 
             m_clphy_entry_train, m_clphy_funding_train,
             m_clpriv_baseline_train, m_clpriv_v2csreprss_train,
             m_clpriv_total_train, m_clpriv_advocacy_train, 
             m_clpriv_entry_train, m_clpriv_funding_train))
})

@@

@@

>>>>

  # Load actual model objects
  tar_load(c(m_pts_baseline_train, m_pts_v2csreprss_train,
             m_pts_total_train, m_pts_advocacy_train, 
             m_pts_entry_train, m_pts_funding_train,
             m_pts_baseline_rewb_train, m_pts_v2csreprss_rewb_train, 
             m_clphy_baseline_train, m_clphy_v2csreprss_train, 
             m_clphy_total_train, m_clphy_advocacy_train, 
             m_clphy_entry_train, m_clphy_funding_train,
             m_clpriv_baseline_train, m_clpriv_v2csreprss_train,
             m_clpriv_total_train, m_clpriv_advocacy_train, 
             m_clpriv_entry_train, m_clpriv_funding_train,
             m_lhr_baseline_train, m_lhr_v2csreprss_train,
             m_lhr_total_train, m_lhr_advocacy_train, 
             m_lhr_entry_train, m_lhr_funding_train))
})

@@

# E~1d~: NGO laws and latent human rights

```{r e1d-predictions, warning=FALSE}
seed_lhr_prediction <- 3207  # From random.org

set.seed(seed_lhr_prediction)
pred_lhr_baseline <- predict(m_lhr_baseline_train, 
                             newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                             allow_new_levels = TRUE)

set.seed(seed_lhr_prediction)
pred_lhr_total <- predict(m_lhr_total_train, 
                          newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                          allow_new_levels = TRUE)

set.seed(seed_lhr_prediction)
pred_lhr_advocacy <- predict(m_lhr_advocacy_train, 
                             newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                             allow_new_levels = TRUE)

set.seed(seed_lhr_prediction)
pred_lhr_entry <- predict(m_lhr_entry_train, 
                          newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                          allow_new_levels = TRUE)

set.seed(seed_lhr_prediction)
pred_lhr_funding <- predict(m_lhr_funding_train, 
                            newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                            allow_new_levels = TRUE)

pred_lhr_baseline_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_baseline)) %>% 
  filter(!is.na(latent_hr_mean_lead1))

pred_lhr_total_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_total)) %>% 
  filter(!is.na(latent_hr_mean_lead1))

pred_lhr_advocacy_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_advocacy)) %>% 
  filter(!is.na(latent_hr_mean_lead1))

pred_lhr_entry_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_entry)) %>% 
  filter(!is.na(latent_hr_mean_lead1))

pred_lhr_funding_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_funding)) %>% 
  filter(!is.na(latent_hr_mean_lead1))
```

```{r e1d-pred-plots, warning=FALSE}
plot_baseline <- ggplot(pred_lhr_baseline_test, 
                        aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Baseline",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_total <- ggplot(pred_lhr_total_test, 
                     aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Total legal barriers",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_advocacy <- ggplot(pred_lhr_advocacy_test, 
                        aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to advocacy",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_entry <- ggplot(pred_lhr_entry_test, 
                     aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to entry",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_funding <- ggplot(pred_lhr_funding_test, 
                       aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to funding",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

```{r e1d-metrics}
lotsa_metrics <- metric_set(rmse, rsq, mae, ccc)

bind_rows(lotsa_metrics(pred_lhr_baseline_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Baseline"),
          lotsa_metrics(pred_lhr_total_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Total legal barriers"),
          lotsa_metrics(pred_lhr_advocacy_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Barriers to advocacy"),
          lotsa_metrics(pred_lhr_entry_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Barriers to entry"),
          lotsa_metrics(pred_lhr_funding_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Barriers to funding")) %>% 
  select(Model, .metric, .estimate) %>% 
  pivot_wider(names_from = ".metric", values_from = ".estimate")
```

```{r e1d-most-improved-total}
most_improved_lhr_total <- bind_cols(
  pred_lhr_baseline_test %>% 
    mutate(error_baseline = Estimate - latent_hr_mean_lead1) %>% 
    select(country, gwcode, year, latent_hr_mean_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_lhr_total_test %>% 
    mutate(error_total = Estimate - latent_hr_mean_lead1) %>% 
    select(estimate_total = Estimate, error_total)
) %>% 
  mutate(improved = error_total < 0 & abs(error_total) < abs(error_baseline)) %>% 
  mutate(diff = error_total - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_lhr_total)

most_improved_lhr_total %>% 
  select(Country = country, Year = year, 
         `Actual latent human rights value (t + 1)` = latent_hr_mean_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_total,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

```{r e1d-most-improved-advocacy}
most_improved_lhr_advocacy <- bind_cols(
  pred_lhr_baseline_test %>% 
    mutate(error_baseline = Estimate - latent_hr_mean_lead1) %>% 
    select(country, gwcode, year, latent_hr_mean_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_lhr_advocacy_test %>% 
    mutate(error_advocacy = Estimate - latent_hr_mean_lead1) %>% 
    select(estimate_advocacy = Estimate, error_advocacy)
) %>% 
  mutate(improved = error_advocacy < 0 & abs(error_advocacy) < abs(error_baseline)) %>% 
  mutate(diff = error_advocacy - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_lhr_advocacy)

most_improved_lhr_advocacy %>% 
  select(Country = country, Year = year, 
         `Actual latent human rights value (t + 1)` = latent_hr_mean_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_advocacy,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

```{r e1d-most-improved-entry}
most_improved_lhr_entry <- bind_cols(
  pred_lhr_baseline_test %>% 
    mutate(error_baseline = Estimate - latent_hr_mean_lead1) %>% 
    select(country, gwcode, year, latent_hr_mean_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_lhr_entry_test %>% 
    mutate(error_entry = Estimate - latent_hr_mean_lead1) %>% 
    select(estimate_entry = Estimate, error_entry)
) %>% 
  mutate(improved = error_entry < 0 & abs(error_entry) < abs(error_baseline)) %>% 
  mutate(diff = error_entry - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_lhr_entry)

most_improved_lhr_entry %>% 
  select(Country = country, Year = year, 
         `Actual latent human rights value (t + 1)` = latent_hr_mean_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_entry,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

```{r e1d-most-improved-funding}
most_improved_lhr_funding <- bind_cols(
  pred_lhr_baseline_test %>% 
    mutate(error_baseline = Estimate - latent_hr_mean_lead1) %>% 
    select(country, gwcode, year, latent_hr_mean_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_lhr_funding_test %>% 
    mutate(error_funding = Estimate - latent_hr_mean_lead1) %>% 
    select(estimate_funding = Estimate, error_funding)
) %>% 
  mutate(improved = error_funding < 0 & abs(error_funding) < abs(error_baseline)) %>% 
  mutate(diff = error_funding - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_lhr_funding)

most_improved_lhr_funding %>% 
  select(Country = country, Year = year, 
         `Actual latent human rights value (t + 1)` = latent_hr_mean_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_funding,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

@@

# E~2d~: Civil society environment and latent human rights

```{r e2d-predictions, warning=FALSE}
seed_lhr_prediction <- 3207  # From random.org

set.seed(seed_lhr_prediction)
pred_lhr_baseline <- predict(m_lhr_baseline_train, 
                                newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                                allow_new_levels = TRUE)

set.seed(seed_lhr_prediction)
pred_lhr_v2csreprss <- predict(m_lhr_v2csreprss_train, 
                                  newdata = select(canary_testing_lagged, -latent_hr_mean_lead1),
                                  allow_new_levels = TRUE)

pred_lhr_baseline_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_baseline)) %>% 
  filter(!is.na(latent_hr_mean_lead1))

pred_lhr_v2csreprss_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, latent_hr_mean_lead1) %>% 
  bind_cols(as_tibble(pred_lhr_v2csreprss)) %>% 
  filter(!is.na(latent_hr_mean_lead1))
```

```{r e2d-pred-plots, warning=FALSE}
plot_baseline <- ggplot(pred_lhr_baseline_test, 
                        aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Baseline",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_v2csreprss <- ggplot(pred_lhr_v2csreprss_test, 
                          aes(x = latent_hr_mean_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Added predictor",
       x = "Actual latent human rights value (t + 1)",
       y = "Predicted latent human rights value (t + 1)") +
  coord_equal() +
  theme_ngo()

```{r e2d-metrics}
lotsa_metrics <- metric_set(rmse, rsq, mae, ccc)

bind_rows(lotsa_metrics(pred_lhr_baseline_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Baseline"),
          lotsa_metrics(pred_lhr_v2csreprss_test, 
                        truth = latent_hr_mean_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Added predictor")) %>% 
  select(Model, .metric, .estimate) %>% 
  pivot_wider(names_from = ".metric", values_from = ".estimate")
```

```{r e2d-most-improved}
most_improved_lhr <- bind_cols(
  pred_lhr_baseline_test %>% 
    mutate(error_baseline = Estimate - latent_hr_mean_lead1) %>% 
    select(country, gwcode, year, latent_hr_mean_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_lhr_v2csreprss_test %>% 
    mutate(error_v2csreprss = Estimate - latent_hr_mean_lead1) %>% 
    select(estimate_v2csreprss = Estimate, error_v2csreprss)
) %>% 
  # mutate(across(starts_with("error"), ~round(., 4))) %>% 
  mutate(improved = error_v2csreprss < 0 & abs(error_v2csreprss) < abs(error_baseline)) %>% 
  mutate(diff = error_v2csreprss - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_lhr)

most_improved_lhr %>% 
  select(Country = country, Year = year, 
         `Actual latent human rights value (t + 1)` = latent_hr_mean_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_v2csreprss,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```
